### 2023-10-25 state/ai.tex
Renamed files and folder
Corrected state of the art

\acf{AI} is very popular, and is getting more attention, being used as selling point by lots of services, the word is thrown around so much that it has lost its meaning. So what are \acp{AI} exactly.

\ac{AI} is defined as theories and stragies used to simulate human intelligence, in other words, a computer or software that can take a decision by itself based on previously set rules. The definition is very wide and have lots of ways to be implemented.

The most promising type of \ac{AI} is machine learning, which consists of teaching the software to make decision. It is defined as the use and development of computer systems that are able to learn and adapt without following explicit instructions, by using algorithms and statistical models to analyse and draw inferences from patterns in data.

@@

\begin{figure}[H]
  \centering
  \includesvg[width=\textwidth]{ml.svg}
  \caption{Machine learning and its subsets}
  \label{fig:ml}
\end{figure}

>>>>

\acf{AI} is very popular, and is getting more attention, being used as selling point by lots of services, the word is thrown around so much that it has lost its meaning. So let us see what are \acp{AI} exactly.

\ac{AI} is defined as theories and strategies used to simulate human intelligence, in other words, a computer or software that can take a decision by itself based on previously set rules. The definition is very wide and has lots of ways to be implemented.

The most promising type of \ac{AI} is machine learning, which in a broad sense means teaching the software to make decisions. It is defined as the use and development of computer systems that are able to learn and adapt without following explicit instructions, by using algorithms and statistical models to analyse and draw inferences from patterns in data.

@@

\begin{figure}[H]
  \centering
  \includesvg[width=\textwidth,pretex=\small]{ml.svg}
  \caption{Machine learning and its subsets}
  \label{fig:ml}
\end{figure}

### 2023-10-25 state/crossbar.tex
Renamed files and folder
Corrected state of the art

Setting memristors in a crossbar array allows to perform analog \ac{VMM} also called Multiply and Accumulate. This circuit works great for this use as all the computation is done almost instantly and using physics laws. Figure \ref{fig:crossbar} shows what a typical crossbar array looks like.

@@

It uses physical properties of electrical systems to perform analog computing. Lets use the circuit node in figure \ref{fig:crossNode} to explain the theory behind the memristor crossbar array.
\begin{figure}[H]
  \centering
  \includesvg[height=3.5cm]{crossbar/node}
  %\def\svgheigth{3.5cm}
  %\input{crossbar/node.pdf_tex}
  \caption{Memristor crossbar node of the $i^{th}$ line and $j^{th}$ column}
  \label{fig:crossNode}
\end{figure}

First of all, a voltage is applied on the $i^{th}$ line, because every column is virtually grounded, the voltage applied to the memristor, of a set resistance $R_i$, is $V_i$. As such and using Ohm's law, we know the the current flowing into the memristor ($I_{mem}$) is bound by the following equation :

@@

This line then joins the column where a current of $I_{i,j-1}$ is flowing, then according to kirchhoff's current law the resulting current is :
\begin{equation}
  I_{j,i} = I_{j,i-1}+I_{mem} = I_{j,i-1} + V_i\cdot\sigma_i
\end{equation}
By applying this process to the whole system we find that the current as the bottom of one column is :
\begin{equation}
  I_1=\sigma_1\cdot V_1 + \sigma_2\cdot V_2 + \sigma_3\cdot V_3
\end{equation}
With $\sigma_1$, $\sigma_2$ and $\sigma_3$ being the conductance of the 3 memristors in the first column.

@@

\begin{itemize}
  \item This circuit being an analog it allows for analog computations which are much faster than their digital counter parts. As explained in this section, the circuit only uses physical properties of the different components and can compute large \acp{VMM} in a very short amount of time.
  \item The other advantage of such a system is the use of the memristors, that make the need to copy data from another memory useless as the memristors have a long term memory themselves. This is one of the ways to break down the Von Neumann bottleneck.
\end{itemize}

>>>>

Setting memristors in a crossbar array allows to perform analog \ac{VMM}, also called Multiply and Accumulate. This circuit works great for this use as all the computation is done almost instantly and using physics laws. Figure \ref{fig:crossbar} shows what a typical crossbar array looks like.

@@

It uses physical properties of electrical systems to perform analog computing. In what follows I will use the circuit node in figure \ref{fig:crossNode} to explain the theory behind the memristor crossbar array.
\begin{figure}[H]
  \centering
  \includesvg[height=3.5cm]{crossbar/node}
  %\def\svgheigth{3.5cm}
  %\input{crossbar/node.pdf_tex}
  \caption{Memristor crossbar node of the $i^{th}$ line and $j^{th}$ column}
  \label{fig:crossNode}
\end{figure}

First of all, a voltage is applied on the $i^{th}$ line, because every column is virtually grounded, the voltage applied to the memristor, of a set resistance $R_i$, is $V_i$. As such and using Ohm's law, we know that the current flowing into the memristor ($I_{mem}$) is bound by the following equation :

@@

This line then joins the column where a current of $I_{i,j-1}$ is flowing, then according to Kirchhoff's current law the resulting current is :
\begin{equation}
  I_{j,i} = I_{j,i-1}+I_{mem} = I_{j,i-1} + V_i\cdot\sigma_i
\end{equation}
By applying this process to the whole system we find that the current at the bottom of one column is :
\begin{equation}
  I_1=\sigma_1\cdot V_1 + \sigma_2\cdot V_2 + \sigma_3\cdot V_3
\end{equation}
With $\sigma_1$, $\sigma_2$ and $\sigma_3$ being the conductance of the 3 memristors in the first column.

@@

\begin{itemize}
  \item By being analog, this circuit allows for analog computations which are much faster than their digital counterparts. As explained in this section, the circuit only uses physical properties of the different components and can compute large \acp{VMM} in a very short amount of time.
  \item The other advantage of such a system is the use of the memristors, that make the need to copy data from another memory useless as the memristors have a long term memory themselves. This is one of the ways to break down the Von Neumann bottleneck.
\end{itemize}

### 2023-10-25 state/gru.tex
Corrected state of the art

The \ac{GRU} is very often compared to the \ac{LSTM}, it is sometimes assimilated as a type of \ac{LSTM} \cite{nbLSTM}. Their performance was found to be very similar in most situations, making those two types of \acp{RNN} coexistant in the modern machine learning world.

There are two versions of the \ac{GRU}, both are found on the internet, they are known as the encoder and decoder version \cite{gru}. They were originally designed to encode the message to translate and then decode in the translation. PyTorch only supports the decoder version \cite{gruPyTorch}, while the keras library supports both \cite{gruKeras} chosen by changing an argument.

@@

\begin{figure}[H]
  \centering
  \includesvg[width=\textwidth,pretex=\large]{gru/encoderCell.svg}
  \label{fig:lstmCell}
  \caption{Encoder \acs{GRU} cell, legend in \cref{leg:cells}}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includesvg[width=\textwidth,pretex=\large]{gru/decoderCell.svg}
  \label{fig:lstmCell}
  \caption{Decoder \acs{GRU} cell, legend in \cref{leg:cells}}
\end{figure}

>>>>

The \ac{GRU} is very often compared to the \ac{LSTM}, being sometimes assimilated as a type of \ac{LSTM} \cite{nbLSTM}. Their performance was found to be very similar in most situations, making those two types of \acp{RNN} coexistant in the modern machine learning world.

There are two versions of the \ac{GRU}, both are found on the internet, they are known as the encoder and decoder version \cite{gru}. They were originally designed to encode the message to translate and then decode in the translation. PyTorch only supports the decoder version \cite{gruPyTorch}, while the Keras library supports both \cite{gruKeras} chosen by changing an argument.

@@

A visual representation of the encoder \ac{GRU} cell is available in \cref{fig:encoderGruCell}.

\begin{figure}[H]
  \centering
  \includesvg[width=\textwidth,pretex=\large]{gru/encoderCell.svg}
  \label{fig:encoderGruCell}
  \caption{Encoder \acs{GRU} cell, legend in \cref{leg:cells}}
\end{figure}

@@

A visual representation of the decoder \ac{GRU} cell can be found in \cref{fig:decoderGruCell}.

\begin{figure}[H]
  \centering
  \includesvg[width=\textwidth,pretex=\large]{gru/decoderCell.svg}
  \label{fig:decoderGruCell}
  \caption{Decoder \acs{GRU} cell, legend in \cref{leg:cells}}
\end{figure}

### 2023-10-25 state/lstm.tex
Corrected state of the art

\section{\acs{LSTM}}\label{sec:lstm}
\acfp{LSTM} are a type of \ac{RNN} used to analyze sequence of data. They are capable of predict data based of the previous results. \acp{LSTM} are part of \acp{RNN}. \acp{RNN} are different than feedfoward \acl{NN} because of their feedback connections. Meaning that the results from the last time step have an impact on the next time step.

@@

\acp{LSTM} was created to fix the vanishing gradient problem. \acp{LSTM} alleviate this issue by adding a cell state, this state gives it the ability to choose what input is important and which one isn't, thus giving it a long term memory. This ability gave the uncommon name of \acl{LSTM} as it has both long and short term memory. This what gives \acp{LSTM} their use for sequence data. They can analyze the data and keep the information from the last time step to make a better decision afterwards. The most comprehensible example is considering a sentence. (TODO : find example)

An \ac{LSTM} is more complicated than just a simple feedforward \acl{NN}, they have several gates, which are all technically a \ac{NN} themselves. There is also a cell state which job is to hold a value for the next step.

@@

\Cref{fig:lstmCell} shows the complexity of the \ac{LSTM} architecture. In an \ac{LSTM}, each gate is a different \ac{NN} and then activated with either a \ac{tanh} or a sigmoid activation function. Each input to the cell is a vector.
Those vector are of varying sizes depending on several factors. For example, both $h_t$ and $c_t$ are of the same size as the number of hidden state (sometimes refered to as cell state) for any $t\geq 0$.
The input vector ($x_t$) is of size of the sample we want to feed for each time step.

@@

The equations of an LSTM are quite unusual.
Let's start with the more classic gate equations. They are the ones that behave like the more classic \ac{NN}.
The input (\cref{eq:inputG}), forget (\cref{eq:forgetG}) and ouput (\cref{eq:outputG}) gates are described below.

@@

Where ($w_i$,$\overrightarrow{b_i}$), ($w_f$,$\overrightarrow{b_f}$) and ($w_o$,$\overrightarrow{b_o}$) are the pair of weights matrixes and biases vectors for the input, forget and ouput gates respectively. $\overrightarrow{x_t}$ is the input vector and $\overrightarrow{h_t}$ is the hidden state vector.

For the next equation describes the candidate cell state (\cref{eq:candCell}), that will next become the cell state (\cref{eq:cellS}).

@@

Using \ac{LSTM} can be a bit tricky. Due to it's sequential nature, it takes several time steps. Every hidden state ($h_t$) is passed to the next time step as an input. This hidden state can be used to compute an output at any time step $t$.

@@

\acp{LSTM} come in reality in a few different flavor. Usually, when \acp{LSTM} are mentionned, the version used is the \ac{NP} version. This is the most common version as those are the equations described on the wikipedia page \cite{wikiLSTM} and used in librairies like Keras \cite{Keras} or pyTorch \cite{PyTorch}. In fact, there are total of 8 other variations of \acp{LSTM} \cite{nbLSTM}.
The difference between each of them varies from one to the other, but they can all be described with new equations.

\subsubsection{Vanilla \ac{LSTM}}
This variant of the \ac{LSTM} was origanally the only \ac{LSTM} network, hence it's name. It differs from the classic \ac{LSTM} with it's use of peephole weights, hence the name of the classic \ac{LSTM} being \acl{NP}. The Vanilla \ac{LSTM} is thus defined by those equations \cite{vanillaLSTM, nbLSTM} :

@@

With $R_{fi}$ being the weight matrix of the input feedback connection to compute the forget gate, and the same goes for the other weight matrices.

>>>>

\section{\acs{LSTM}}\label{sec:lstm}
\acfp{LSTM} are a type of \ac{RNN} used to analyze sequences of data. They are capable of predicting data based on previous data points.

@@

\acp{LSTM} was created to fix the vanishing gradient problem. \acp{LSTM} alleviate this issue by adding a cell state, this state gives it the ability to choose what input is important and which one is not, thus giving it a long term memory. This ability gave the uncommon name of \acl{LSTM} as it has both long and short term memory. This is what makes \acp{LSTM} adequate for sequence data. They can analyze the data and keep the information from the last time step to make a better decision afterwards. The most comprehensible example is considering a sentence. %TODO : find example)

An \ac{LSTM} is more complicated than just a simple feedforward \acl{NN}, it has several gates, which all compute a \ac{VMM}. There is also a cell state whose job is to hold a value for the next step.

@@

\Cref{fig:lstmCell} shows the complexity of the \ac{LSTM} architecture. In an \ac{LSTM}, each gate is a different \ac{NN} and then activated with either a \ac{tanh} or a sigmoid activation function. Each input to the cell is a vector.
Those vectors are of varying sizes depending on several factors. For example, both $h_t$ and $c_t$ are of the same size as the number of hidden states (sometimes referred to as cell state) for any $t\geq 0$.
The input vector ($x_t$) is of size of the sample we want to feed for each time step.

@@

The equations of an LSTM are quite unusual.
Let's start with the more classic gate equations. They are the ones that behave like the more classic \ac{NN}.
The input (\cref{eq:inputG}), forget (\cref{eq:forgetG}) and output (\cref{eq:outputG}) gates are described below.

@@

Where ($w_i$,$\overrightarrow{b_i}$), ($w_f$,$\overrightarrow{b_f}$) and ($w_o$,$\overrightarrow{b_o}$) are the pair of weights matrixes and biases vectors for the input, forget and output gates respectively. $\overrightarrow{x_t}$ is the input vector and $\overrightarrow{h_t}$ is the hidden state vector.

The next equation describes the candidate cell state (\cref{eq:candCell}), that will next be used to compute the cell state (\cref{eq:cellS}).

@@

Using \ac{LSTM} can be a bit tricky. Due to its sequential nature, it takes several time steps. Every hidden state ($h_t$) is passed to the next time step as an input. This hidden state can be used to compute an output at any time step $t$. \Cref{fig:lstmUse} shows a visual representation of an \ac{LSTM} going from the current time step to the following time step.

@@

\acp{LSTM} come in a few different flavors of implementations. Usually, when \acp{LSTM} are mentionned, the version used is the \ac{NP} version. This is the most common version as those are the equations described on the wikipedia page \cite{wikiLSTM} and used in libraries like Keras \cite{Keras} or PyTorch \cite{PyTorch}. In fact, there are at least 8 other variations of \acp{LSTM} \cite{nbLSTM}.
The difference between them varies from one to the other, some of them are detailled next.

\subsubsection{Vanilla \ac{LSTM}}
This variant of the \ac{LSTM} was originally the only \ac{LSTM} network, hence its name. It differs from the classic \ac{LSTM} with its use of peephole weights, hence the name of the classic \ac{LSTM} being \acl{NP}. The Vanilla \ac{LSTM} is thus defined by the following equations \cite{vanillaLSTM, nbLSTM} :

@@

With $R_{fi}$ being the weight matrix of the input feedback connection to compute the forget gate, and the same goes for the other weight matrixes.

### 2023-10-25 state/memristors.tex
Renamed files and folder
Corrected state of the art

Memristors are the lesser known fourth fundamental passive component of electronics, among resistors, capacitors and inductor.
It was first theorized in 1971 by L. Chua from UC Berkeley, in \cite{TheoMemristor}. The name comes from the blend of \textit{memory} and \textit{resistance}.
The theory behind this component was extracted from a missing component to link the four fundamental circuit variables, voltage ($v$), charge ($q$), current ($i$) and flux ($\phi$). \Cref{fig:fundComp} shows the four fundamental variables are on each side of the square, with the ones on opposite sides being linked by the following equations :
\begin{equation}
  d\phi = v\cdot dt
\end{equation}

@@

Resistors, capacitors and inductors were already very established and well known component, so it was theorized that a fourth device should then exists to physically link flux ($\phi$) and charge ($q$).  The flux in this case is not a magnetic flux and is defined as such : $ d\phi=v\cdot dt \Rightarrow \phi =  \int v \,dt  $.\\
The component stayed theoretical until 2008 when it's been implemented in a physical device for the first time \cite{memristorFab}. It took 37 years to have an actual working device.\\
There is then an extention of this theoretical device to another, the memristive device. It was theorized in 1976 by L. Chua and S. M. Kang \cite{memrestiveDev}. The difference between the memristor and the memristive device is its internal behavior. Memristive device are commonly referred to as memristor as well.

@@

\subsection{Equations}
A memristor links the flux ($\phi$) and charge ($q$) and creates the memristance.
This memristance is defined with the following equation :
\begin{equation}
  M(q)=\frac{d\phi}{dq}
\end{equation}
It can be compared with the other fundamental components like resistor ($R(i)=\frac{dv}{di}$), capacitor ($\frac{1}{C(q)}=\frac{dv}{dq}$) and inductor ($L(i)=\frac{d\phi}{di}$).
We can then extract a more useful equation in an actual circuit :
\begin{equation}
  v(t)=M(q(t))\cdot i(t)
\end{equation}
Similarly, a memductance can be defined as such :
\begin{equation}
  W(\phi)=\frac{dq}{d\phi}
\end{equation}
A memristive device is slightly differently defined, it still uses a memristance, but here the memristance also depends on an internal state called $x$. This gives us this equation :
\begin{equation}
  v(t)=M(x,i)\cdot i(t)
\end{equation}
The internal state ($x$) is not linked to flux or charge in the case of a memristive device.\\
Once again, we can also define the memristive device using a memductance :
\begin{equation}
  i(t)=W(x,v)\cdot v(t)
\end{equation}
In all of the previous equations, $v$ is the voltage in Volt ($V$), $i$ is the current in Ampere ($A$), $\phi$ is the flux in Weber ($Wb$), $q$ is the charge in Coulomb ($C$), $M$ is the memristance in Ohm ($\Omega$) and $W$ is the memductance in Siesmens ($S$ or $\Omega^{-1}$).

@@

A memristor is defined as a non-linear two-terminal fundamental electrical component. It behaves as a resistance with memory (hence its name), meaning that it changes its resistance based on how much charge went through it. This enables us to manipulate the resistance of the component.
The huge benefit of memristors its ability to retain its internal resistance, the device can left without power for a long period of time (retention time of minimum 10 years according to \cite{memRetention}). When the device is powered backup, it will have the same resistance it had before.

@@

\subsection{Usage}
The main research for memristor usage is using them as ReRAM. The idea behind ReRAM is to use memristors as Non-Volatile memory. It uses two states of the device with known resistances ($R_{on}$ and $R_{off}$), giving it binary property. Reading the memory simply requires setting a voltage and reading the output current. It is better than current solutions (HDD, SSD) as it has a much lower latency. It is better than traditionnal DRAM because it keeps the information even when turned off. This makes ReRAM a good replacement for both RAM and HDDs/SSDs, thus eliminating Von Neumann bottleneck due to the Von Neumann architecture that is used in all modern computers.

>>>>

Memristors are the lesser known fourth fundamental passive component of electronics, among resistors, capacitors and inductor.
It was first theorized in 1971 by L. Chua from UC Berkeley, in \cite{TheoMemristor}. The name comes from the blend of \textit{memory} and \textit{resistance}.
The theory behind this component was extracted from a missing component to link the four fundamental circuit variables, voltage ($v$), charge ($q$), current ($i$) and flux ($\phi$). \Cref{fig:fundComp} shows the four fundamental variables are on each side of the square, with the ones on opposite sides being linked by the following equations :

\begin{equation}
  d\phi = v\cdot dt
\end{equation}

@@

Resistors, capacitors and inductors were already very established and well known components, so it was theorized that a fourth device should then exist to physically link flux ($\phi$) and charge ($q$).  The flux in this case is not a magnetic flux and is defined as such : $ d\phi=v\cdot dt \Rightarrow \phi =  \int v \,dt  $.

The component stayed theoretical until 2008 when it was implemented in a physical device for the first time \cite{memristorFab}. It took 37 years to have an actual working device.

There is then an extension of this theoretical device to another, the memristive device. It was theorized in 1976 by L. Chua and S. M. Kang \cite{memrestiveDev}. The difference between the memristor and the memristive devices is its internal behavior. Memristive device are commonly referred to as memristors as well.

@@

\subsection{Equations}
A memristor links the flux ($\phi$) and charge ($q$) by the memristance.
This memristance is defined with the following equation :
\begin{equation}
  M(q)=\frac{d\phi}{dq}
\end{equation}
It can be compared with the other fundamental components like resistor ($R(i)=\frac{dv}{di}$), capacitor ($\frac{1}{C(q)}=\frac{dv}{dq}$) and inductor ($L(i)=\frac{d\phi}{di}$).
We can then extract a more useful equation in an actual circuit :
\begin{equation}
  v(t)=M(q(t))\cdot i(t)
\end{equation}
Similarly, a memductance can be defined as such :
\begin{equation}
  W(\phi)=\frac{dq}{d\phi}
\end{equation}
A memristive device is slightly differently defined, it still uses a memristance, but here the memristance also depends on an internal state called $x$. This gives us this equation :
\begin{equation}\label{eq:memristiveDev}
  v(t)=M(x,i)\cdot i(t)
\end{equation}
The internal state ($x$) is not linked to flux or charge in the case of a memristive device.\\
Once again, we can also define the memristive device using a memductance :
\begin{equation}
  i(t)=W(x,v)\cdot v(t)
\end{equation}
In all of the previous equations, $v$ is the voltage in Volt ($V$), $i$ is the current in Ampere ($A$), $\phi$ is the flux in Weber ($Wb$), $q$ is the charge in Coulomb ($C$), $M$ is the memristance in Ohm ($\Omega$) and $W$ is the memductance in Siemens ($S$ or $\Omega^{-1}$).

@@

A memristor is defined as a non-linear two-terminal fundamental electrical component. It behaves as a resistance with memory (hence its name), meaning that it changes its resistance based on how much charge went through it. This enables us to manipulate the resistance of the component.
The huge benefit of memristors is the ability to retain its internal resistance, the device can be left without power for a long period of time (retention time of minimum 10 years according to \cite{memRetention}). When the device is powered backup, it will have the same resistance it had before.

@@

\subsection{Usage}
The main research for memristor usage is using them as ReRAM. The idea behind ReRAM is to use memristors as Non-Volatile memory. It uses two states of the device with known resistances ($R_{on}$ and $R_{off}$), giving it binary property. Reading the memory simply requires setting a voltage and reading the output current. It is better than current solutions (HDD, SSD) as it has a much lower latency. It is better than traditionnal DRAM because it keeps the information even when turned off. This makes ReRAM a good replacement for both RAM and HDDs/SSDs, thus eliminating von Neumann bottleneck due to the von Neumann architecture that is used in all modern computers.

### 2023-10-25 state/model.tex
Corrected state of the art

It's been established that the internal resistance also depends on an internal state $x$ (\cref{eq:memristiveDev}).

@@

\subsection{Behavior}

The model was tested then tested. As with physical memristors, it has a threshold voltage value from which its resistive states starts changing. This threshold value is $V_{read}$. For this specific model the threshold voltage is set to $V_{read}=500mV$.

Every pulse is composed of 2 parts and last for a total of $1 ms$ :
\begin{itemize}
  \item The reading of the resistive state, by applying a triangular wave that in bound by the threshold value $V_{read}$.
  \item The writing to the memristor. During this part the resistive state will change. This is done by apply a square wave of variable width ($t_{\Delta}$) with a set bias voltage ($V_{bias}$).
\end{itemize}

\begin{figure}[H]
  \centering
  \includesvg[width=\textwidth]{memristor/input.svg}
  \label{graph:memInput}
  \caption{Pulse shape graph}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includesvg[width=\textwidth]{memristor/memPulseChange.svg}
  \label{graph:pulseChange}
  \caption{Memristor's resistive state under different pulses width (Fixed bias voltage $V_{bias}=1.8V$)}
\end{figure}

The effect of the pulse width variation is shown in \cref{graph:pulseChange}. Indeed, the larger the pulse widths, the faster the internal resistance reaches the bounding resistance (\cref{eq:memModel4}). The bounding resistance being the same as all curve use the same bias voltage ($V_{bias}=1.8V$), the curves will all eventually reach the max resistance.

\begin{figure}[H]
  \centering
  \includesvg[width=\textwidth]{memristor/memVoltChange.svg}
  \label{graph:voltChange}
  \caption{Memristor's resistive state under different voltages (Fixed pulse width $t_{\Delta}=100\mu s$)}
\end{figure}

@@

The resistive state can also be lowered by applying negative voltages.TODO

>>>>

It has been established that the internal resistance also depends on an internal state $x$ (\cref{eq:memristiveDev}).

@@

\subsection{Behavior}\label{subsec:memModelBehave}

The model implements, as with physical memristors, a threshold voltage value from which its resistive states starts changing. This threshold value is $V_{read}$. For this specific model the threshold voltage is set to $V_{read}=500mV$.

Every pulse is composed of 2 parts and lasts for a total of $1 ms$ :
\begin{itemize}
  \item The reading of the resistive state, by applying a triangular wave that is bound by the threshold value $V_{read}$. The triangular wave starts from $0V$, goes to $V_{read}$ then to $-V_{read}$ before going back to $0V$.
  \item The writing to the memristor. During this part the resistive state will change. This is done by applying a square wave of variable width ($t_{\Delta}$) with a set bias voltage ($V_{bias}$).
\end{itemize}

\begin{figure}[H]
  \centering
  \includesvg[width=\textwidth]{memristor/input.svg}
  \caption{Pulse shape graph}
  \label{graph:memInput}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includesvg[width=\textwidth]{memristor/memPulseChange.svg}
  \caption{Memristor's resistive state under different pulses width (Fixed bias voltage $V_{bias}=1.8V$)}
  \label{graph:pulseChange}
\end{figure}

The effect of the pulse width variation is shown in \cref{graph:pulseChange}. Indeed, the larger the pulse widths, the faster the internal resistance reaches the bounding resistance (\cref{eq:memModel4}). The bounding resistance being the same as all curves use the same bias voltage ($V_{bias}=1.8V$), the curves will all eventually reach the max resistance.

\begin{figure}[H]
  \centering
  \includesvg[width=\textwidth]{memristor/memVoltChange.svg}
  \caption{Memristor's resistive state under different voltages (Fixed pulse width $t_{\Delta}=100\mu s$)}
  \label{graph:voltChange}
\end{figure}

@@

The resistive state can also be lowered by applying negative voltages.%TODO

### 2023-10-25 state/nn.tex
Corrected state of the art

The model presented in \cref{fig:snn} is modular and can be scaled up as much as required. It is generally accepted that the more neurons a \ac{NN} has, the best it can learn and execute the given task.

@@

The weights values are obtained through training. This is done in a few steps :

\begin{enumerate}
  \item Generate random values for each weight.
  \item Run the \ac{NN} to get an output vector.\label{step:restart}
  \item Measure error rate using the chosen error algorithm.
  \item Use the
  \item go back to \cref{step:restart}
\end{enumerate}

>>>>

The model presented in \cref{fig:snn} is modular and can be scaled up as much as required. It is generally accepted that the more neurons a \ac{NN} has, the more complex the problems  it can solve can be.

@@

The weights values are obtained through training. In supervised learning, in order train the weights it is required to have a dataset of inputs/outputs values that are correct. The outputs are the target values, the values that the \ac{NN} needs to compute when being executed. The training starts once the weights have been initialized (usually randomly picked values). The training is an iteravtive process, each iteration being called an epoch. Each of those epoch usally consists of the following steps :

\begin{enumerate}
  \item Run the \ac{NN} to get an output vector.\label{step:restart}
  \item Measure the error (known as loss) by comparing the current prediction with the targeted output using the chosen error algorithm (\ac{RMSE}, \ac{MSE}, etc).
  \item Run the backpropagation algorithm to change each individual weight.
\end{enumerate}

The number of epochs required depends on the complexity of the problem the \ac{NN} is solving.

Once trained, the \ac{NN} can be tested by feeding the \ac{NN} input data it hasn't seen before.

### 2023-10-25 state/rnn.tex
Corrected state of the art

\acp{RNN} are, as the name suggests, a type of \ac{NN} using recurrent connections. They are a \ac{NN} with at least one cycle within the structure, where outputs of the previous run is used the next one. Those feedback connections are what differenciates it from feedforward \ac{NN}.

@@

Traditionnal \ac{RNN} have the ability to model sequential events by propagating through time, for example forward and backward propagation. This is achieved by connecting these sequential events with the hidden state like in \cref{eq:rnn}.

@@

They are trained the same way \ac{NN} are, measuring the error, backpropagte and adjust the weights accordingly.

@@

The Vanishing gradient problem is a problem that comes when dealing with time depedent data \cite{vanishGrad}. When big amount of time depedent data is fed to the \ac{RNN}, the weights can't be updated properly. The older the data, the lower it will impact how much the weight must change. Rendering the old input data almost useless. Simple \acp{RNN} must be used with relatively short time series.

Some \acp{RNN} were designed around this issue. This is the case of the \ac{LSTM} and \ac{GRU} which were created with internal mechanisms to regulate the flow of information and gradients.

>>>>

\acp{RNN} are, as the name suggests, a type of \ac{NN} using recurrent connections. They are a \ac{NN} with at least one cycle within the structure, outputs of previous time step are used as input for the next time step, these outputs are generally known are hidden states. Those feedback connections are the main difference from feedforward \ac{NN}.

@@

Traditionnal \acp{RNN} have the ability to model sequential events by propagating through time, for example forward and backward propagation. This is achieved by connecting these sequential events with the hidden state like in \cref{eq:rnn}.

@@

They are trained the same way \ac{NN} are, measuring the error, backpropagate and adjust the weights accordingly.

@@

The Vanishing gradient problem is a problem that comes when dealing with time dependent data \cite{vanishGrad}. When a big amount of time dependent data is fed to the \ac{RNN}, the weights cannot be updated properly. The older the data, the lower it will impact how much the weight must change. Rendering the old input data almost useless. That is the reasom why, simple \acp{RNN} must be used with relatively short time series.

Some \acp{RNN} were designed to tackle this issue. This is the case of the \ac{LSTM} and \ac{GRU} which were created with internal mechanisms to regulate the flow of information and gradients.
